\documentclass[11pt]{article}

% --- packages ---
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools}
\usepackage{booktabs}
\usepackage{array}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{tcolorbox}

% --- theorem environments ---
\newtheorem{theorem}{Theorem}
\theoremstyle{definition}
\newtheorem{remark}{Remark}

% --- callout boxes ---
\newtcolorbox{verifybox}{colback=green!4,colframe=green!45!black,boxrule=0.6pt,
  left=6pt,right=6pt,top=4pt,bottom=4pt,arc=2pt}
\newtcolorbox{caveatbox}{colback=yellow!8,colframe=orange!70!black,boxrule=0.6pt,
  left=6pt,right=6pt,top=4pt,bottom=4pt,arc=2pt}
\newtcolorbox{brokenbox}{colback=red!5,colframe=red!60!black,boxrule=0.7pt,
  left=6pt,right=6pt,top=4pt,bottom=4pt,arc=2pt}

% --- shorthands ---
\newcommand{\R}{\mathbb{R}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\Szero}{\Sigma_0}
\newcommand{\PM}{P_{\!M}}
\newcommand{\PN}{P_{\!N}}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator*{\Real}{Re}

\title{\textbf{$\Sigma_0$ --- The Collapse Certificate}\\[4pt]
\large A computable stability certificate for convergence dynamics, and an honest
account of why an ungrounded self-improving system tends to collapse or diverge}
\author{Lantern OS \\ \texttt{src/cio\_sde/}}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
We give a computable certificate for contraction in a dissipative nonlinear
system: splitting the local Jacobian's symmetric part into active and near-null
subspaces yields a Lyapunov function $V(x)=\tfrac12\norm{\PM x}^2$ whose decay is
guaranteed when the active spectral abscissa is negative \emph{and the active
subspace is invariant under the full Jacobian} (e.g.\ $A$ normal). We are explicit
about that hypothesis: the general non-normal case is not covered by the energy
argument, and the authoritative test is the full spectrum $\max\Real\lambda(A)<0$.
For the non-normal regime we add two sufficient, provable contraction gates (a
numerical-range gate and a Lyapunov/similarity-metric gate, \S\ref{sec:gates768})
that strictly widen the certified region.
We pair the theorem with a semantic collapse operator $\Szero$ (a heuristic
trigger plus projection onto the null manifold --- the ``42-state''), an
anti-collapse operator $\Sigma_0^{-1}$ (empirically shown, not proven, to suppress
collapse), and an early-warning signal. We read the construction as a small, honest
warning about recursively self-improving systems: absent external grounding, the
linearized dynamics admit only collapse or divergence. Theorem~1 is machine-checked
($31$ passing, \#506 resolved) for the symmetric case, with $9$ further tests for the
non-normal gates (\texttt{test\_stability\_gates.py}). The other components are heuristics,
labelled as such; one referenced demonstration is not currently reproducible (its
scripts are absent from the repository); the AI-safety reading is the keeper and
stands on its own.
\end{abstract}

\begin{caveatbox}
\textbf{Per-section status.} \S1 theorem: \emph{proven under an explicit hypothesis,
machine-checked for the symmetric case}. \S2 trigger: \emph{operational definition,
not derived}. \S3 anti-collapse: \emph{heuristic, $N{=}1$ empirical}. \S4 canary:
\emph{the eigenvalue readout is unimplemented; the correct signal (innovation
surprise) is implemented in \texttt{surprise.py}}. \S5 graph: \emph{standard}.
\S6 demonstration: \emph{not reproducible --- scripts absent}. \S7 AI-safety:
\emph{the keeper}.
\end{caveatbox}

\tableofcontents
\bigskip

% ====================================================================
\section{The object}
We study a dissipative nonlinear system
\begin{equation}
\dot{x} = f(x,u,\theta), \qquad x \in \R^{n},
\end{equation}
with internal state $x$ (for the router: a conversation's encoded state),
control / persistent-excitation input $u$, and slowly-varying parameters $\theta$.
Linearizing along a trajectory $x^\ast$ gives the local Jacobian
\begin{equation}
\dot{\delta x} = A\,\delta x, \qquad
A = \left.\frac{\partial f}{\partial x}\right|_{x^\ast},
\end{equation}
and everything below reasons about the eigenstructure of $A$ and its symmetric
part $A_s = \tfrac12(A + A^{\top})$. The skew part $A_k=\tfrac12(A-A^\top)$ carries
rotation and is \emph{not} seen by $A_s$; the gap between the two is exactly what
makes the general case in \S2 harder than the symmetric one.

% ====================================================================
\section{The collapse-guarantee theorem}
Split the state space using $A_s$ into the \emph{near-null} subspace
$N = \operatorname{span}\{\, v_i : |\lambda_i(A_s)| < \varepsilon \,\}$ and its
orthogonal complement, the \emph{active} subspace $M$ with projector $\PM$. Define
the Lyapunov function on the active modes only,
\begin{equation}
V(x) = \tfrac12 \norm{\PM x}^2,
\end{equation}
and let the active spectral abscissa be
$\alpha = \max\{\, \lambda_i(A_s) : v_i \in M \,\}$.

\begin{theorem}[Contraction on the active subspace]
Assume $\alpha < 0$ \textbf{and} the active subspace $M$ is $A$-invariant, i.e.
$\PM A \PN = 0$ (equivalently: $A$ is normal, $A=A_s$, or $A$ commutes with $\PM$).
Then
\begin{equation}
\dot V \le 2\alpha V
\quad\Longrightarrow\quad
\norm{\PM x(t)} \le \norm{\PM x(0)}\, e^{\alpha t}.
\end{equation}
The active modes decay exponentially at rate $|\alpha|$ and the trajectory
contracts onto the invariant null manifold $N$: under this hypothesis, collapse is
guaranteed. If $\alpha \ge 0$, some active mode is non-contracting and collapse is
not guaranteed.
\end{theorem}

\begin{proof}
Differentiating $V$ along the flow,
\[
\dot V = (\PM x)^{\top}\PM A\,x
= (\PM x)^{\top} A_s (\PM x)
\;+\; \underbrace{(\PM x)^{\top} A\,(\PN x)}_{\text{cross term}}.
\]
The first term is $\le 2\alpha V$ by the definition of $\alpha$. The cross term
vanishes exactly when $M$ is $A$-invariant ($\PM A \PN = 0$), which holds
automatically if $A$ is normal. Under that hypothesis $\dot V \le 2\alpha V$, and
Gr\"onwall gives the exponential bound.
\end{proof}

\subsection{Why the hypothesis is required}
\label{sec:hypothesis}
For a general non-normal $A$ the skew part $A_k$ couples the active and null
components, the cross term need not vanish, and the simple energy bound can fail.

\begin{caveatbox}
\textbf{Counterexample.} For $A=\begin{psmallmatrix}-1 & 3\\ -3 & 0\end{psmallmatrix}$
the active abscissa is $\alpha=-1<0$, yet a direct scan finds
$\max \dot V/V \approx +8.8\times10^{4} \gg 2\alpha=-2$, and integrating from
$x_0=(0.3,1.0)$ makes $V$ \emph{grow} $0.045\to0.341$ --- a sign violation of
$\dot V\le 2\alpha V$. Collapse still occurs here, but it is rescued by a different
argument, not by the energy proof. The earlier ``exact, not approximate'' wording
applies only to the symmetric case.
\end{caveatbox}

\subsection{The authoritative test: the full spectrum}
$\alpha<0$ on the symmetric part is \emph{necessary but not sufficient} for strict
contraction of the full system. The conservative, always-correct condition is
\begin{equation}
\max \Real\,\lambda(A) < 0 \quad\text{on the \emph{full} } A
\ (\text{via \texttt{eig}, not \texttt{eigvalsh}}).
\end{equation}
A Bendixson bound gives $\Real\lambda(A)\le\lambda_{\max}(A_s)$, so $\alpha$ bounds
the real parts but does not by itself certify them: a rotation such as
$A=\begin{psmallmatrix}-1&0&0\\0&0&2\\0&-2&0\end{psmallmatrix}$ has $\alpha=-1$ yet
eigenvalues $\{-1,\pm 2i\}$, a center that never collapses in its rotating plane.
\textbf{Recommended:} report both $\alpha$ and $\max\Real\lambda(A)$. As of \#505,
\texttt{collapse\_certificate()} reports both (the conservative small-gain $\alpha$ and
the exact $\max\Real\lambda(A)$); \S\ref{sec:gates768} adds two provable gates that
widen the certified non-normal region.

\begin{verifybox}
\textbf{Verification.} The shipped test uses $A=-0.8\,I$, which is \emph{symmetric}
(the hypothesis holds exactly, cross term identically zero). The certificate
predicts contraction rate $0.8$ and a rollout shows $V$ decaying monotonically.
This confirms the theorem in precisely the special case where the proof is
unconditionally valid; it does \emph{not} exercise the non-normal case.
(\texttt{collapse\_certificate}, \texttt{lyapunov\_value}, \texttt{src/cio\_sde/collapse.py}.)
\end{verifybox}

\subsection{Provable region-wideners for non-normal $A$ (\#768)}
\label{sec:gates768}
The small-gain bound over-rejects strongly non-normal $A$, and $\max\Real\lambda(A)<0$
alone is necessary-not-sufficient (transient growth). \texttt{stability\_gates()} adds
two \emph{sufficient, provable} contraction certificates, each strictly wider than
small-gain (margin $m\ge0$):
\begin{enumerate}
\item \textbf{Numerical-range gate (monotone).}
  $\omega(A)=\lambda_{\max}(A_s)<-m \implies \norm{e^{tA}}_2\le e^{\omega t}$ --- a
  strict, no-transient contraction (matrix measure $\mu_2$; Lohmiller--Slotine 1998).
  $\omega(A)$ is the rightmost point of the numerical range $W(A)$; since
  $\operatorname{spec}(A)\subset W(A)$, this gate \emph{implies} the Lyapunov gate.
\item \textbf{Lyapunov gate (asymptotic, optimal metric).} For margin $m\ge0$,
  $\exists P\succ0:\ (A+mI)^\top P+P(A+mI)\prec0 \iff \max\Real\lambda(A)<-m$ (classical
  Lyapunov theorem; $=\inf_T\mu_2(TAT^{-1})<-m$). Certified by solving
  $(A+mI)^\top P+P(A+mI)=-I$ and checking $P\succ0$ (with a relative ill-conditioning
  guard, so the boundary is never certified independent of solver warnings). It
  \emph{accepts the \S\ref{sec:hypothesis}~counterexample}
  $A=\begin{psmallmatrix}-1&3\\-3&0\end{psmallmatrix}$ (Hurwitz at $-\tfrac12$) that the
  energy proof and the small-gain bound reject; $\sqrt{\kappa(P_0)}$ (the margin-0 solve)
  bounds the Euclidean transient $\sup_t\norm{e^{tA}}$.
\end{enumerate}
\textbf{Honest scope.} Sufficient, not necessary; they certify the \emph{full}
Jacobian's contraction, not collapse-onto-manifold (the active/null alignment gap is
separate). The \emph{proven} transient constant is Crouzeix--Palencia (2017):
$\norm{e^{tA}}\le 1+\sqrt2$ when $W(A)\subset$~LHP ($\omega\le0$); the sharper constant
$2$ is Crouzeix's still-open conjecture.
\begin{verifybox}
\textbf{Verification.} \texttt{tests/test\_stability\_gates.py} (9 tests): the
$\begin{psmallmatrix}-1&3\\-3&0\end{psmallmatrix}$ case (small-gain rejects, Lyapunov
accepts, numerical-range flags the transient), a 400-matrix red-team finding \emph{no}
false-positive certificate, and matrix-exponential checks that the monotone ($e^{\omega t}$),
Lyapunov ($\sqrt{\kappa(P)}$), and Crouzeix--Palencia ($1+\sqrt2$) bounds each hold.
(\texttt{stability\_gates}, \texttt{StabilityGates}, \texttt{src/cio\_sde/collapse.py}.)
\end{verifybox}

% ====================================================================
\section{The collapse trigger $\Sigma_0$}
\label{sec:sigma0}
\begin{caveatbox}
\textbf{Status: operational definition --- not derived from Theorem~1.}
\end{caveatbox}
$\Szero$ fires when \emph{all four} conditions hold (a control singularity over a
structureless field):
\begin{center}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}>{$}l<{$}l@{}}
\toprule
\textrm{Condition} & \textrm{Meaning} \\
\midrule
\norm{\nabla_x L} < \varepsilon_g & no optimization signal remains \\
\rank(J_f) < \rho\, n & drift Jacobian has lost directional structure \\
\Sigma \textrm{ isotropically flat} & uncertainty has no preferred direction \\
\norm{\partial H/\partial u} < \varepsilon_c & control cannot distinguish actions \\
\bottomrule
\end{tabular}
\end{center}
None of these four quantities is the spectral abscissa $\alpha$ that Theorem~1 uses.
They are an operational definition of ``underdetermined'' --- a soft AND-gate
$p=\min(p_\nabla,p_{\text{rank}},p_{\text{aniso}},p_{\text{ctrl}})$ over four
independent signs of degeneracy. The link between ``the four conditions fire'' and
``$\alpha<0$'' is a modeling assumption, not a proof. When triggered, $\Szero$
projects the state onto the null eigenmodes of $A_s$,
\begin{equation}
x^\ast = P x, \qquad P = V_{\mathrm{null}} V_{\mathrm{null}}^{\top},
\end{equation}
the ``42-state'' (a colloquial name, no formal meaning).

\begin{caveatbox}
\textbf{Caveat.} $x^\ast=Px$ is a true fixed point only when $A$ is normal. For
non-normal $A$ the projection kills the symmetric part but the skew rotation leaves
$\norm{A x^\ast}$ large, so $x^\ast$ is not an equilibrium of the real flow. In the
implementation the apparent ``freeze'' is produced by the integrator overwriting
$x_{\text{next}}=x^\ast$ and discarding the diffusion term --- a clamp, not an
emergent attractor. (\texttt{SemanticCollapseOperator}.)
\end{caveatbox}

% ====================================================================
\section{The anti-collapse operator $\Sigma_0^{-1}$}
\begin{caveatbox}
\textbf{Status: control-design heuristic. Empirically shown ($N{=}1$) to suppress
collapse. No sufficiency theorem --- unlike Theorem~1.}
\end{caveatbox}
Where $\Szero$ projects \emph{onto} the null subspace, $\Sigma_0^{-1}$ injects
energy \emph{along} it:
\begin{equation}
dx = f\,dt + dW + \Sigma_0^{-1}, \qquad
\Sigma_0^{-1} = s\cdot p \cdot (V_{\mathrm{null}}\,\xi),
\end{equation}
with $\xi$ standard normal, $s$ the strength, $p\in[0,1]$ the collapse proximity.
The implementation correctly injects noise in the null subspace and
re-anisotropizes $\Sigma$. There is, however, \emph{no companion theorem} stating
that $\Sigma_0^{-1}$ prevents collapse --- only a single forced-collapse
demonstration. The classical persistent-excitation result (Anderson 1977) concerns
estimator convergence; here there is no estimator, so the PE invocation is
\emph{analogy only}.

\begin{verifybox}
\textbf{Evidence ($N{=}1$).} On one forced-collapse run, $\Szero$ fired every step
and the state froze; with $\Sigma_0^{-1}$ active, $\Szero$ stopped firing and the
state escaped. Encouraging directional evidence, not a guarantee. (The often-quoted
``$40/40\to0/40$'' and ``$0.05\to12.9$'' figures are illustrative log values, not
pinned by any test; the forced-collapse tests run $20$--$30$ steps.)
\end{verifybox}

\begin{remark}
Latent defect (fixed): \texttt{AntiCollapseOperator.\_\_init\_\_} annotated
\texttt{detector: Optional[...]} while \texttt{collapse.py} imported only
\texttt{Dict}; the module loaded only because PEP~563 keeps annotations as strings.
Corrected to \texttt{from typing import Dict, Optional}.
\end{remark}

% ====================================================================
\section{The early-warning scalar (the ``canary'')}
\begin{caveatbox}
\textbf{Status: the eigenvalue readout below is NOT implemented in code. The right
signal --- innovation surprise --- now is (\texttt{src/cio\_sde/surprise.py}).}
\end{caveatbox}
Near a bifurcation the dominant eigenvalue flattens (critical slowing down;
\cite{wissel1984,scheffer2009}). Two \emph{proposed} readouts,
$p_{\mathrm{unbounded}} = 1/|\Real\lambda_{\max}(A_s)|$ and
$p_{\mathrm{gate}} = \operatorname{clip}(1-|\Real\lambda_{\max}|/\varepsilon,0,1)$,
are written here but appear nowhere in \texttt{collapse.py}; the actual driver of
$\Sigma_0^{-1}$ is the four-condition $\texttt{proximity()}$, not $p_{\mathrm{gate}}$.

\begin{verifybox}
\textbf{The right canary, implemented.} The correct early-warning is \emph{surprise
relative to uncertainty}: the Kalman normalized innovation squared
$\mathrm{NIS}=\nu^\top S^{-1}\nu$ with $\nu=y-C\hat x$, $S=C\Sigma C^\top + R$.
$\mathrm{NIS}\approx m$ means model and reality agree; $\mathrm{NIS}\gg m$ means the
model is overconfident and has drifted without knowing it. This is the standard
innovation-consistency $\chi^2$ test \cite{barshalom2001}, implemented in
\texttt{SurpriseMonitor}; \texttt{experiments/sigma0\_horse\_blinders.py} shows a
low-coupling observer staying \emph{calm while wrong} before the surprise spikes.
The dangerous state is the quiet before the spook, not the spook.
\end{verifybox}

% ====================================================================
\section{Global structure: the attractor graph $G$}
The system is multistable. Collect its attractors $\{A_1,\dots,A_k\}$, each with a
basin $B_i = \{\, x_0 : \lim_{t\to\infty}\phi_t(x_0) \in A_i \,\}$. Coarse-graining
to a graph $G=(V,E)$ gives an induced Markov process over basins,
\begin{equation}
P_{ij}(u) = \Pr\!\big(\pi(x_{t+1}) = A_j \mid \pi(x_t) = A_i,\, u_t\big),
\end{equation}
with partition map $\pi:\R^n\to V$. This is the standard Markov-state-model
construction; it is genuinely Markov only under timescale separation (fast
intra-basin relaxation vs.\ slow inter-basin hops). A boundary studded with
\emph{saddles} has stable manifolds --- ridges one can ride with control.
\texttt{experiments/sigma0\_saddle\_ride.py} measures the cost: spin (the imaginary
part $\Im\lambda$) is orthogonal to capture (the real part), and only structured
control --- which knows the unstable direction and spends energy --- rides the
ridge. \texttt{experiments/sigma0\_boundary\_qutrit.py} maps such a boundary from
outside as pure negative space, verified against an analytic separatrix.

% ====================================================================
\section{Demonstration on router data}
\label{sec:demo}
\begin{brokenbox}
\textbf{Reproducibility flag --- read first.} This section describes a pipeline
whose two driver scripts, \texttt{experiments/router\_sigma0\_encoder.py} and
\texttt{experiments/router\_reservoir\_G.py}, are \textbf{absent from the
repository} (verified across all branches). No numeric result below has a committed
generating artifact. Treat this section as a planned / not-yet-committed sketch, not
an established result.
\end{brokenbox}
The intended demonstration encodes each conversation turn as
$x=[\text{novelty},\text{self\_repeat},\text{echo},\text{length}]\in[0,1]^4$ and
runs the construction two ways (local Jacobian fitting; reservoir reconstruction).
The narrative result --- a \emph{parrot attractor}
($\text{novelty}\approx1,\text{echo}\approx0.72$), i.e.\ model collapse in the data
--- is \textbf{withdrawn} as unverified: because the generating scripts do not
exist, those numbers have no produced artifact, and the claim that the result
``appears directly in the data, not inserted by hand'' cannot be substantiated.
\emph{To make \S6 real:} commit the two scripts with a logged run, then replace the
withdrawn claims with the produced numbers.

% ====================================================================
\section{Why this is a warning against ASI}
\begin{verifybox}
\textbf{Status: the one downstream claim worth keeping --- as a machine-learning
claim, on ML evidence, not as a consequence of the physics or of \S6.}
\end{verifybox}
A system that optimizes against its own representations with no external anchor is
the flow $\dot x = f(x)$ where $f$ only ever sees $x$. The linearized certificate
frames two degenerate fates absent outside contact: \textbf{collapse} onto a dead
fixed point (the 42-state, mirrors agreeing with mirrors), or \textbf{divergence}.
The only stable middle required an external bound. \textbf{Grounding is the safety
mechanism.} This does not need the certificate's physics or the \S6 numbers: it is
the documented phenomenon of \emph{model collapse} \cite{shumailov2024}, closely
related to reward hacking \cite{amodei2016}. The ``parrot attractor'' is model
collapse renamed. Two honest qualifications: the strict ``collapse or diverge, no
third'' dichotomy is the \emph{linearized} framing (real training also admits limit
cycles and partial-information equilibria --- better stated as ``tends to degenerate
or destabilize absent grounding''); and the in-repo demonstration inherits \S6's
missing-scripts problem, so the \emph{argument} rests on the published literature,
not the demo.

% ====================================================================
\begin{thebibliography}{9}
\bibitem{lyapunov1892} A.\,M. Lyapunov.
  \emph{The General Problem of the Stability of Motion}. 1892.
\bibitem{poincare} H. Poincar\'e.
  \emph{M\'emoire sur les courbes d\'efinies par une \'equation diff\'erentielle}. 1880s.
\bibitem{bendixson1901} I. Bendixson.
  Sur les courbes d\'efinies par des \'equations diff\'erentielles. \emph{Acta Math.}, 1901.
\bibitem{wissel1984} C. Wissel.
  A universal law of the characteristic return time near thresholds. \emph{Oecologia}, 1984.
\bibitem{scheffer2009} M. Scheffer et al.
  Early-warning signals for critical transitions. \emph{Nature}, 2009.
\bibitem{anderson1977} B.\,D.\,O. Anderson.
  Exponential stability of linear equations arising in adaptive identification. \emph{IEEE TAC}, 1977.
\bibitem{barshalom2001} Y. Bar-Shalom, X.-R. Li, T. Kirubarajan.
  \emph{Estimation with Applications to Tracking and Navigation}. Wiley, 2001.
\bibitem{pathak2018} J. Pathak et al.
  Model-free prediction of large spatiotemporally chaotic systems from data. \emph{PRL}, 2018.
\bibitem{shumailov2024} I. Shumailov et al.
  AI models collapse when trained on recursively generated data. \emph{Nature}, 2024.
\bibitem{amodei2016} D. Amodei et al.
  Concrete Problems in AI Safety. \emph{arXiv:1606.06565}, 2016.
\end{thebibliography}

\begin{caveatbox}
Bibliography verified 2026-06-17 (issue \#660): all entries are established
classical references (Lyapunov 1892, Poincar\'e, Bendixson 1901, Anderson 1977
IEEE TAC, Bar-Shalom 2001, Pathak 2018 PRL, Shumailov 2024 Nature, Amodei 2016
arXiv:1606.06565). The companion \texttt{.md} additionally cites the verified
model-collapse scaling-law results (Dohmatob et al.\ arXiv:2402.07043; Feng et al.\
arXiv:2406.07515).
\end{caveatbox}

\paragraph{Implementation and testing.}
\begin{itemize}
\item \texttt{src/cio\_sde/collapse.py} --- \texttt{SemanticCollapseOperator},
  \texttt{CollapseCertificate}/\texttt{collapse\_certificate}, \texttt{AntiCollapseOperator},
  \texttt{ReconstructionOperator}, \texttt{stability\_gates}/\texttt{StabilityGates} (\#768).
\item \texttt{src/cio\_sde/surprise.py} --- \texttt{SurpriseMonitor} and the Kalman
  measurement update (the missing inference step).
\item \texttt{tests/test\_cio\_sde.py} --- 31 passing (\#506 resolved); Theorem~1
  exercised in the symmetric case only.
\item \texttt{tests/test\_stability\_gates.py} --- 9 passing: the \#768 numerical-range
  $+$ Lyapunov gates, a 400-matrix red-team (no false-positive certificate), and
  matrix-exponential checks of the proven transient bounds.
\item Runnable, deterministic experiments present in the repo:
  \texttt{sigma0\_reconstruction.py} (recoverability floor),
  \texttt{sigma0\_horse\_blinders.py} (the surprise canary),
  \texttt{sigma0\_saddle\_ride.py} (spin $\perp$ capture),
  \texttt{sigma0\_boundary\_qutrit.py} (mapping a wall by negative space).
\item \textbf{Absent:} \texttt{router\_sigma0\_encoder.py},
  \texttt{router\_reservoir\_G.py} --- \S\ref{sec:demo} is not reproducible until
  they are committed.
\end{itemize}

\end{document}
